Multiresolution Neural Field Estimation

### 2011-03-01 Multiresolution_framework/ltx/manuscript.tex
minor edits
trivial edits

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Multiresolution Neural Field Estimation }

\author{In no particular order: \parham{P. Aram}, \mike{M. Dewar}, \dean{D.R. Freestone,~\IEEEmembership{Graduate Student Member,~IEEE,}}, \ken{K.Scerri}, D.B. Grayden,~\IEEEmembership{Member,~IEEE,} and V. Kadirkamanathan % <-this % stops a space
\thanks{P. Aram* and V. Kadirkamanathan are with the Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, S1 3JD, U.K. (e-mail:p.aram@sheffield.ac.uk; visakan@sheffield.ac.uk).}% <-this % stops a space
\thanks{M. Dewar is with the School of Informatics, University of Edinburgh, Edinburgh, EH8 9AB, U.K. (e-mail:mike.dewar@inf.ed.ac.uk)}
\thanks{D.\ R.\ Freestone and D.\ B.\ Grayden are with the Department
of Electrical and Electronic Engineering, The University of Melbourne, and The Bionic Ear Institute, VIC, Australia,  {\tt\small dfreestone@bionicear.org, grayden@unimelb.edu.au}}}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

@@

@@

\begin{abstract}
%\boldmath
The Integro-difference equation (IDE) is an increasingly popular model of spatio-temporal processes. Here we develop a multiresolution approximation (MRA) framework for the IDE neural field equations based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is approached in a Maximum Likelihood (ML) framework using the Expectation Maximisation (EM) algorithm. Examples are given to demonstrate the framework.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Neural field model, multiresolution approximation (MRA), Expectation Maximization (EM) algorithm, wavelets.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Introduction}
\IEEEPARstart{T}{he} human brain has a multiresolution architecture, where spatial scales for information processing and transfer range from individual proteins in synapses, to neurons, to networks of neurons. The multiresolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multiresolution data-driven framework for neural field modeling, the multiresolution approximate integrodifference equation (MRAIDE), to address this challenge.

Neural field models typically describe the meso and macroscopic dynamics of the human brain, but are also descriptive of finer-scale neurodynamics. The ability to create patient-specific neural models will contribute to our knowledge base of poorly understood diseases such as epilepsy, and furthermore, may also enable the development of new treatment strategies. This is particularly poignant with the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using systems theory. The design of a stimulation strategy using systems theory requires a model. Neural mass and neural field models are the ideal candidates for estimation and control algorithms. It is expected that parameters of the neural fields models will be patient specific and will therefore need to be inferred from data.

The first work describing data driven mesoscopic neural modeling~\cite{Valdes1999} used a neural mass model~\cite{LopesDaSilva1976,Zetterberg1978} to fit EEG data. This approach was extended to coupled neural masses (using another model~\cite{Jansen1995}) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM)~\cite{David2003}. Following this work, data-driven modeling was extended to continuum field equations that explain the richer dynamics of spatiotemporal neural fields \cite{Galka2008,schiff2008kalman,Daunizeau2009}. Most recently a framework was developed where a finite element model of the neural field (via a global Galerkin projection) was formed, using a basis function decomposition, to transform the PDE neural field equations into a finite dimension system to facilitate efficient state and parameter estimation~\cite{Freestone2011}.

This paper is an extension to this recent study~\cite{Freestone2011}. The basis function decomposition in this previous study did not account for multiresolution architecture and spatial dynamics of the human cortex. Thus, it can be improved.. In this current paper, a multiresolution approach is taken...

@@

\section{Integro-Difference Equation Neural Field Model}
The stochastic integro-difference equation (IDE) form of the Amari neural field  formulation~\cite{Amari1977} is given by (see~\cite{Freestone2011} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The term $e_t(\mathbf r)$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$, where $\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$  denotes a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. For simplicity, \cut{we denote the current time step by $t$ and the future time step by $t+1$.} The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f\left(.\right)$. \dean{The activation function is of the form $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r}) + v_0$}. A linear form of the activation function was chosen to construct the estimator due to high computational demands when using a nonlinear (sigmoidal) activation function.

The observation equation describing the electrophysiological recordings is given by 
\begin{equation}\label{eq:ObservationEquation}
	\mathbf y_t = \int_{\Omega} { m\left(\mathbf{r}_{n}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t, 
\end{equation}
where $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$, is corrupted by an i.i.d normally distributed zero-mean white noise $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$ with $\mathbf{\Sigma}_{\epsilon}=\sigma_{\epsilon}^2\mathbf I_{n_y} $. The superscript $\top$ denotes the transpose operator.
\section{MRA of the IDE in State-Space}
The multiresolution approximation (MRA) of the neural IDE is obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a one-dimensional field, the connectivity kernel is decomposed by,
\begin{equation}
 w\left(r-r'\right)=\sum_{l \in \mathbb{Z}}\alpha_{j_0,l}\phi_{j_0,l}\left(r-r'\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l}\psi_{j,l}\left(r-r'\right), 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with $\phi_{j,l}\left(r\right)=2^{\frac{j}{2}}\phi\left(2^jr-l\right) $ and $\psi_{j,l}\left(r\right)=2^{\frac{j}{2}}\psi\left(2^jr-l\right)$. Integers $j$ and $l$ are the scale and translation parameters respectively. The field is decomposed as
\begin{equation}
 v_t\left(r\right)=\sum_{l \in \mathbb{Z}}x_{t,j_{0},l}\phi_{j_{0},l}\left(r\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}} \check{x}_{t,j,l}\psi_{j,l}\left(r\right),
\label{eq:FieldExpansion}
\end{equation}
where $ x_{t,j_{0},l}$ and $\check{x}_{t,j,l} $ are the coefficients of the expansion at time $t$.

Eqs. \eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} are infinite series expansions, but  for a practical implementation they must be truncated at some level $j$. Therefore, the field and the connectivity kernel is approximated by
 \begin{align}
 w\left(r-r'\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right) 
\label{eq:KernelFiniteExpansion} \\
 v_t\left(r\right) &\approx \boldsymbol\mu\left(r\right)^\top\mathbf{x}_t,
\label{eq:FieldFiniteExpansion}
\end{align}
where the unknown parameter and state vectors, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
\begin{align}
 \boldsymbol\theta^\top &=\left[ \boldsymbol\alpha_{j_0}^\top \quad \boldsymbol\beta_{j_0}^\top \quad \boldsymbol\beta_{j_0+1}^\top \cdots \boldsymbol\beta_{j}^\top\right] 
\label{KernelWeights} \\
 \mathbf{x}_{t}^\top &=\left[\mathbf{x}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}+1}^\top \cdots \check{\mathbf{x}}_{t,j}^\top\right].
\label{FieldWeights}
\end{align}
The kernel and field approximation coefficient vectors $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$ contain all the coefficients $\left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace $ and $\left\lbrace x_{t,j_0, l}: l \in \mathbb{Z}\right\rbrace$, respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$ contain all the coefficients $\left\lbrace \beta_{j,l} :l \in \mathbb{Z}\right\rbrace$ and $\left\lbrace  \check x_{t,j, l}:l \in \mathbb{Z}\right\rbrace$ respectively.

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$ respectively, are defined by
\begin{align}
 \boldsymbol\lambda^\top & (r-r')=\left[ \boldsymbol\phi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0+1}^\top(r-r') \right. \nonumber \\
&\left. \cdots \quad \boldsymbol\psi_{j}^\top(r-r')\right] \label{KernelBasisVector} \\
 \boldsymbol\mu^\top & (r)=\left[ \boldsymbol\phi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0+1}^\top(r) \cdots \boldsymbol\psi_{j}^\top(r) \right]. 
\label{FieldBasisVector}
\end{align}
The vectors in Eqs. (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in Eqs. (\ref{KernelWeights}) and (\ref{FieldWeights}).

@@

  \newpage
% \bibliographystyle{unsrt} 
% \bibliography{MRAIDE}

@@

\end{document}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Multiresolution Neural Field Estimation }
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{In no particular order: \parham{P. Aram}, \mike{M. Dewar}, \dean{D.R. Freestone}, \ken{K.Scerri}, D.B. Grayden and V. Kadirkamanathan % <-this % stops a space
\thanks{P. Aram* and V. Kadirkamanathan are with the Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, S1 3JD, U.K. (e-mail:p.aram@sheffield.ac.uk; visakan@sheffield.ac.uk).}% <-this % stops a space
\thanks{M. Dewar is with the School of Informatics, University of Edinburgh, Edinburgh, EH8 9AB, U.K. (e-mail:mike.dewar@inf.ed.ac.uk)}}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

@@

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

@@

\begin{abstract}
%\boldmath
% The Integro Difference Equation (IDE) is an increasingly popular model of spatio-temporal processes. Here we develop a Multiresolution Approximation (MRA) framework for the IDE based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is approached in a Maximum Likelihood (ML) framework using the Expectation Maximisation (EM) algorithm. Examples are given to demonstrate the features of the model.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Neural field model, multiresolution approximation (MRA), expectation maximization (EM) algorithm, wavelets.
\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
%  \begin{center} \bfseries EDICS Category: SSP-IDEN \end{center}
%  \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

We have a clear definition and where to stimulate and how to stimulate.\\
\cut{This color is used to indicate bits we can get rid of if we dont have enough space, use the command: \textbackslash cut$\left\lbrace  \right\rbrace$}

\section{Introduction}
\IEEEPARstart{T}{his} paper introduces a multiresolution data-driven framework for neural field modeling. Neural field models describe the meso and macroscopic dynamics of the human brain. Modeling the brain at this level is extremely important, since diseases such as epilepsy and \parham{Parkinson's disease} are manifested at this scale through pathological oscillations. The ability to create patient-specific neural field models will not only contribute to our knowledge base of poorly understood diseases such as epilepsy, but may also enable the development of new treatment strategies. This is particularly poignant with the advent of new devices that use therapeutic electrical stimulation. Presently, stimulation strategies for devices using corrective or modulating electrical stimulation operate in an open-loop, where stimulation parameters are chosen using a process of trial and error. Therefore, there exists an enormous potential to improve the performance of these devices using systems theory. The design of a stimulation strategy using systems theory requires a model. Neural mass and neural field models are the ideal candidates for estimation and control algorithms. It is expected that parameters of the neural fields models will be patient specific and will therefore need to be inferred from data.

Creating patient specific neural models has been approached from a number of different standpoints. \dean{need to add proper citations here.} Valdes et al. (1999) used the neural mass model of Lopes Da Silva (1976) and Zetterberg et al. (1978) to fit EEG data. This approach was extended by David and Friston (2003) to coupled neural masses (using the equations of Jansen and Ritt (1995)) through a Bayesian estimation scheme dubbed Dynamic Causal Modeling (DCM). Following this work, Galka et al. (2008), Sauer et al. (2008), Daunizeau et al. (2009), and Freestone et al (2011) developed the estimation frameworks for continuum neural field equations that explain the richer dynamics of spatiotemporal neural fields.

This paper is an extension to the framework presented Freestone et al. (2011), where a finite element model of the neural field (via a global Galerkin projection) was formed, using a basis function decomposition, to transform the PDE neural field equations into a finite dimension system to facilitate state and parameter estimation. The basis function decomposition in this previous study did not account for all of the spatial dynamics. Thus, it can be improved.. In this current paper, a multiresolution approach is taken...

@@

\section{Integro-Difference Equation Neural Field Model}
The stochastic integro-difference equation (IDE) form of the Amari neural field  formulation~\cite{Amari1977} is given by (see \dean{cite NeuroImage paper} for a full derivation)
\begin{equation}\label{eq:DiscreteTimeModel}
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r'}\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\, d\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where the post-synaptic membrane voltage at time $t$ of a population of neurons at position $\mathbf r$ is denoted $v_t\left(\mathbf r\right)$. Synaptic dynamics are included through the parameter $\xi=\left(1-\frac{ T_s}{\tau}\right)$, where $\tau$ is the synaptic time constant and $T_s$ is the sampling time. The connectivity strength between neural populations at a distance $\mid\mathbf{r}-\mathbf{r'}\mid$ is described by the kernel $w\left(\mathbf{r}-\mathbf{r}'\right)$. The connectivity kernel is taken as a ``Mexican hat'' function, which describes local excitation and lateral inhibition \cite{Amari1977}. The firing rate of the presynaptic neurons is related to the post-synaptic membrane potential by the activation function $f\left(.\right)$. \dean{The activation function is of the form $f(v_t(\mathbf{r})) = \varsigma v_t(\mathbf{r}) + v_0$}. The term $e_t(\mathbf r)$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{s})\sim\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$, where $\mathcal{GP}(\mathbf 0,\eta(\mathbf{r}-\mathbf{r'}))$  denotes a zero mean spatial Gaussian process with covariance function $\eta(\mathbf{r}-\mathbf{r'})$~\cite{Rasmussen2005}. For simplicity, \cut{we denote the current time step by $t$ and the future time step by $t+1$.} The observation equation describing the electrophysiological recordings is given by 
\begin{equation}\label{eq:ObservationEquation}
	\mathbf y_t = \int_{\Omega} { m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right) v_t\left(\mathbf{r}'\right) \, d\mathbf{r}'} + \boldsymbol\epsilon_t, 
\end{equation}
where $\mathbf{y}_{t} = [y_t(\mathbf{r}_1) y_t(\mathbf{r}_2) \cdots y_t(\mathbf{r}_{n_y})]^\top$, compiled at $n_{y}$ spatial locations at time $t$ via the observation kernel $m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)$, is corrupted by an i.i.d normally distributed zero-mean white noise $\boldsymbol{\epsilon}_{t}\sim \mathcal{N}\left(\mathbf{0},\mathbf{\Sigma}_{\epsilon}\right)$ with $\mathbf{\Sigma}_{\epsilon}=\sigma_{\epsilon}^2\mathbf I_{n_y} $. The superscript $\top$ denotes the transpose operator.

\section{MRA of the IDE in State-Space}
The multiresolution approximation (MRA) of the neural IDE is obtained by decomposing both the field, $v_t(.)$, and the connectivity kernel, $w(.)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a one-dimensional field, the connectivity kernel is decomposed by,
 w\left(r-r'\right)=\sum_{l \in \mathbb{Z}}\alpha_{j_0,l}\phi_{j_0,l}\left(r-r'\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l}\psi_{j,l}\left(r-r'\right), 
\label{eq:KernelExpansion}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with $\phi_{j,l}\left(r\right)=2^{\frac{j}{2}}\phi\left(2^jr-l\right) $ and $\psi_{j,l}\left(r\right)=2^{\frac{j}{2}}\psi\left(2^jr-l\right)$. Integers $j$ and $l$ are the scale and translation parameters respectively. The field is decomposed as
 v_t\left(r\right)=\sum_{l \in \mathbb{Z}}x_{t,j_{0},l}\phi_{j_{0},l}\left(r\right)+\sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}} \check{x}_{t,j,l}\psi_{j,l}\left(r\right),
\label{eq:FieldExpansion}
where $ x_{t,j_{0},l}$ and $\check{x}_{t,j,l} $ are the coefficients of the expansion at time $t$.
% A  two-dimensional MRA can be implemented using 2-D scaling and wavelet functions built up  using the tensor-product approach \cite{Meyer1992,Mallat1998}. We write the 2-D scaling function as
% \begin{equation}
%  \phi_{j,\mathbf{l}}\left(r_1,r_2\right)=\phi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right)
% \label{eq:2Dscalingfunction}
% \end{equation}
% and three 2-D wavelet functions $ \psi^{(n)} \left(\mathbf{r}\right)$ where $n=1,2,3$ as
% \begin{equation}
%  \psi_{j,\mathbf{l}}^{(1)}\left(r_1,r_2\right)=\phi_{j,l_1}\left(r_1\right)\psi_{j,l_2}\left(r_2\right)
% \label{eq:2Dwavelet_1}
% \end{equation}
% \begin{equation}
%  \psi_{j,\mathbf{l}}^{(2)}\left(r_1,r_2\right)=\psi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right)
% \label{eq:2Dwavelet_2}
% \end{equation}
% \begin{equation}
%  \psi_{j,\mathbf{l}}^{(3)}\left(r_1,r_2\right)=\psi_{j,l_1}\left(r_1\right)\psi_{j,l_2}\left(r_2\right)
% \label{eq:2Dwavelet_3}
% \end{equation}
% for $j \in \mathbb{Z}$ and, $\mathbf{l} \in \mathbb{Z}^2$, where $\psi_{j,\mathbf{l}}^{(1)},  \psi_{j,\mathbf{l}}^{(2)}$ and $\psi_{j,\mathbf{l}}^{(3)} $ extract fine features of the 2-D field at vertical, horizontal and diagonal orientations respectively.
% This can be generalised to $d$-dimensional MRA by introducing the scaling function
% \setlength{\arraycolsep}{0.0em}
% \begin{eqnarray}
%   \phi_{j,\mathbf{l}}\left(r_1,r_2, \cdots r_d\right)&=&\phi_{j,l_1}\left(r_1\right)\phi_{j,l_2}\left(r_2\right) \cdots \phi_{j,l_d}\left(r_d\right) \label{eq:d-dimensionalscalingfunction}\\
%  &&j \in \mathbb{Z}, \quad \mathbf{l} \in \mathbb{Z}^d \nonumber
% \end{eqnarray}
% \setlength{\arraycolsep}{5pt}
% and $2^d-1$ wavelet functions $ \psi^{(n)} \left(\mathbf{r}\right), n=1, \cdots,2^d-1$ with a procedure similar to the 2-D case. Using these basis functions, multi-dimensional MRA of the kernel and the field can be obtained by
% \setlength{\arraycolsep}{0.0em}
% \begin{eqnarray}
%  k\left(\mathbf{r}-\mathbf{r'}\right)&=&\sum_{\mathbf{l}\in \mathbb{Z}^d}\alpha_{j_0,\mathbf{l}}\phi_{j_0,\mathbf{l}}\left(\mathbf{r}-\mathbf{r'}\right) \nonumber \\
% &+&\sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l}\in \mathbb{Z}^d}\sum_{i=1}^{2^d-1}\beta_{j,\mathbf{l}}^{(i)} \psi_{j,\mathbf{l}}^{(i)}\left(\mathbf{r}-\mathbf{r'}\right) 
% \label{eq:MultidimensionalKernelExpansion}
% \end{eqnarray}
% \setlength{\arraycolsep}{5pt}
% \begin{equation}
% z_t\left(\mathbf{r}\right)=\sum_{\mathbf{l}\in \mathbb{Z}^d}x_{t,j_{0},\mathbf{l}}\phi_{j_{0},\mathbf{l}}\left(\mathbf{r}\right)+\sum_{j\geq j_0}^{\infty} \sum_{\mathbf{l}\in \mathbb{Z}^d}\sum_{i=1}^{2^d-1} \check{x}_{t,j,\mathbf{l}}^{(i)}\psi_{j,\mathbf{l}}^{(i)}\left(\mathbf{r}\right)
% \label{eq:MultidimensionalFieldExpansion}
% \end{equation}
Eqs \eqref{eq:KernelExpansion} and \eqref{eq:FieldExpansion} are infinite series expansions, but  for a practical implementation they must be truncated at some level $j$. Therefore, the field and the connectivity kernel is approximated by
 w\left(r-r'\right) &\approx \boldsymbol\theta^\top\boldsymbol\lambda\left(r-r'\right) 
\label{eq:KernelFiniteExpansion} \\
 v_t\left(r\right) &\approx \boldsymbol\mu\left(r\right)^\top\mathbf{x}_t,
\label{eq:FieldFiniteExpansion}
where the unknown parameter and state vectors \dean{(block matrices?)}, $\boldsymbol\theta \in \mathbb{R}^{n_{\theta}}$ and $\mathbf{x}_t \in \mathbb{R}^{n_x}$, are defined as 
 \boldsymbol\theta^\top &=\left[ \boldsymbol\alpha_{j_0}^\top \quad \boldsymbol\beta_{j_0}^\top \quad \boldsymbol\beta_{j_0+1}^\top \cdots \boldsymbol\beta_{j}^\top\right] 
\label{KernelWeights} \\
 \mathbf{x}_{t}^\top &=\left[\mathbf{x}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}}^\top \quad  \check{\mathbf{x}}_{t,j_{0}+1}^\top \cdots \check{\mathbf{x}}_{t,j}^\top\right].
\label{FieldWeights}
The kernel and field approximation coefficient vectors $\boldsymbol \alpha_{j_0}$ and $\mathbf{x}_{t,j_{0}}$ contain all the coefficients $\left\lbrace\alpha_{j_0, l}:l \in \mathbb{Z} \right\rbrace $ and $\left\lbrace x_{t,j_0, l}: l \in \mathbb{Z}\right\rbrace$ respectively. Similarly, the kernel and the field detail coefficient vectors, $\boldsymbol\beta_{j}$ and $\check{\mathbf{x}}_{t,j}$ contain all the coefficients $\left\lbrace \beta_{j,l} :l \in \mathbb{Z}\right\rbrace$ and $\left\lbrace  \check x_{t,j, l}:l \in \mathbb{Z}\right\rbrace$ respectively.

The vectors \dean{(block matrices?)} of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$ respectively, are defined by
\begin{align}
 \boldsymbol\lambda^\top & (r-r')=\left[ \boldsymbol\phi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0}^\top(r-r') \quad \boldsymbol\psi_{j_0+1}^\top(r-r') \right. \nonumber \\
&\left. \cdots \quad \boldsymbol\psi_{j}^\top(r-r')\right] \label{KernelBasisVector} \\
 \boldsymbol\mu^\top & (r)=\left[ \boldsymbol\phi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0}^\top(r) \quad \boldsymbol\psi_{j_0+1}^\top(r) \cdots \boldsymbol\psi_{j}^\top(r) \right]. 
\label{FieldBasisVector}
\end{align}
The vectors in Eqs. (\ref{KernelBasisVector}) and (\ref{FieldBasisVector}) are constructed in the same manner as the vectors in Eqs. (\ref{KernelWeights}) and (\ref{FieldWeights}).

@@

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
  \newpage
% \bibliographystyle{unsrt} 
% \bibliography{MRAIDE}

@@

% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)

% \begin{thebibliography}{1}
% 
% \bibitem{IEEEhowto:kopka}
% H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
%   0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
% 
%  \end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:
% \begin{IEEEbiography}{Parham Aram}
% 
%  
% \end{IEEEbiography}
% 
% \begin{IEEEbiography}{Visakan Kadirkamanathan}
% % Biography text here.
% (M’90) received the
% B.A. and Ph.D. degrees in electrical and information
% engineering from the University of Cambridge, U.K.
% He held Research Associate positions at the University
% of Surrey, U.K., and the University of Cambridge,
% U.K., before joining the Department of Automatic
% Control and Systems Engineering, The University
% of Sheffield, U.K., as a Lecturer in 1993, where
% he is currently a Professor of Signal and Information
% Processing and is affiliated to the Centre for Signal
% Processing and Complex Systems. His research interests
% include nonlinear signal processing, system identification, intelligent control
% and fault diagnosis with applications in systems biology, aerospace systems,
% and wireless communication. He has coauthored a book on intelligent control
% and has published more than 120 papers in refereed journals and proceedings of
% international conferences.
% Prof. Kadirkamanathan is the Co-Editor of the International Journal of Systems
% Science and has served as an Associate Editor for the IEEE TRANSACTIONS
% ON NEURAL NETWORKS and the IEEE TRANSACTIONS ON SYSTEMS, MAN, AND
% CYBERNETICS, PART B.
%  \end{IEEEbiography}
% \begin{IEEEbiography}{Michael Dewar}
%  received the M.Eng. degree in
% control systems engineering and the Ph.D. degree
% in systems engineering both from The University of
% Sheffield, U.K., in 2002 and 2007, respectively.
%  He is currently working as
% a Research Associate in the Institute for Adaptive
% and Neural Computation, School of Informatics,
% The University of Edinburgh, U.K. 
% \end{IEEEbiography}

% if you will not have a photo at all:
% \begin{IEEEbiographynophoto}{John Doe}
% Biography text here.
% \end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

% \begin{IEEEbiographynophoto}{Jane Doe}
% Biography text here.
% \end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}
